---
title: "Test 1"
author: "Siddhant Raghuvanshi"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries Required
```{r Library to include}
library(rgl)
options(rgl.printRglwidget = TRUE)
```

## Visualization function
```{r Visualization function to plot 2D graphs}
visualize3D <- function(fname ,range){
  
  x<- y<- seq(-range,range,length = 100)
  data<-data.matrix(expand.grid(x,y))

  
  `f(x,y)`<-matrix(apply(data,1,fname),
             nrow = length(x),
             ncol = length(y))
  
  div<-100
  color <- rev(rainbow(div))
  zcol <- cut(`f(x,y)`, div)
  
  persp3d(x,y,`f(x,y)` , col = color[zcol])
}

```

## 1) Ackley Function 

### 1.1) Equation of Ackley function
$$
f(x) = -a\,exp \left(-b\,\sqrt{\frac{1}{d} \sum_{i=1}^{d}{x_i^2}}\right)-
\,exp\left(-b\,{\frac{1}{d} \sum_{i=1}^{d}{cos(cx_i)}}\right) +
\,a+\,exp(1)
$$

### 1.2) R implementation of Ackley function

```{r Ackley function}
ackley<-function(x,a=20, b=0.2 , c =2*pi){
  d <- length(x)
  
  sum1 <- sum(x^2)
  sum2 <- sum(cos(2*pi*x))
  
  term1 <- -a * exp(-b*sqrt(sum1/d))
  term2 <- -exp(sum2/d)
  
  y <- term1 + term2 + a + exp(1)
  return(y)
}

```
### 1.3) Ackley function visualization
```{r Ackley function visualization}
visualize3D(ackley,32)
```


## 2) Zakharov Function

### 2.1) Equation of Zakharov Function

$$
Z_n(x) = \sum_{i=1}^{n}{x_i^2} + \left(\sum_{i=1}^{n}{0.5ix_i}\right)^2 + 
\left(\sum_{i=1}^{n}{0.5ix_i}\right)^4 
$$

### 2.2) R implementation of Zakharov function

```{r zakharov function}
zakharov <- function(x){
  
  term1 <-sum(x^2)
  i <-seq_along(x)
  term2 <- sum(i*0.5*(x))
  
  y <- term1 + term2^2 + term2^4
  return(y)
}
```

### 2.3) Visulization of Zakharov function
```{r Visulization of Zakharov function}
visualize3D(zakharov,10)
```

## 3) Dixon-price function

### 3.1) Dixon-price function equation
$$
f(x) = (x_1 - 1)^2 +\sum_{i=2}^{n}{i(2x_i^2-x_{i-1})^2}
$$

### 3.2) R implementation of Dixon-price function

```{r Dixon-price function}
dixon <- function(x){
  
  term1 <-(x[1]-1)^2
  
  # computation of sum starts from i = 2 thus taking lag befor removing values
  d <- length(x)-1
  xlag <- x[1:d]
  
  x <- x[-1]
  i <- seq_along(x)+1
  
  term2 <- sum(i*(2*x^2 - xlag)^2 ) 
  
  y <- term1 + term2
  
  return(y)
  
}
```
### 3.3) Dixon-price function visulization
```{r Dixon-price function visulization}
visualize3D(dixon,10)
```

## 4) Griewank function
### 4.1) Griewank function equation
$$
f(x) = \sum_{i=n}{n}\frac{x_i^2}{4000}-
\prod_{i=1}^{n}cos\left(\frac{x_i}{\sqrt{i}}\right)+1
$$

### 4.2) R implementation of Griewank function

```{r Griewank function}

griewank <- function(x,y){
  term1 <- sum(x^2)
  term2 <- prod(cos(x)/seq_along(x))
  y <- 1 + (1/4000 * term1) - term2
  
  return(y)
}

```
### 4.3) Visulization of Griewank function
```{r Visulization of Griewank function}
visualize3D(griewank,100)
```

## 5) Michalewicz Function
### 5.1) Michalewicz Function equation
$$
f(x) = - \sum_{i=1}^{d}sin(x_i){sin^{2m}\left(\frac{ix_i^2}{\pi}\right)}
$$

### 5.2) R implementation of Griewank function

```{r Michalewicz Function}
michale <- function(x,m=10){
  
  d <- length(x)
  i <- seq_along(x)
  
  sum <- sum(sin(x) * (sin(i*x^2/pi))^(2*m))
  
  y <- sum
  return(-y)
}

# new
visualize3D(michale,4)
```
## 6) Perm $P_{n,\beta}$ Function
### 6.1) Equation of Perm $P_{n,\beta}$ Function 
$$
f(x)=\sum_{i=1}^{d}{
\left(\sum_{j=1}^{d}{(j^i+\beta)}
  \left({\left({\frac{x_j}{j}}\right)^i}-1\right)
\right)^2}
$$


### 6.2) R implementation of Perm $P_{n,\beta}$ Function 

```{r Prem}

perm <- function(x, b=0.5){
  
  d <- length(x)
  ii <- c(1:d)
  jj <- matrix(rep(ii,times=d), d, d, byrow=TRUE)
  
  xxmat <- matrix(rep(x,times=d), d, d, byrow=TRUE)
  inner <- rowSums((jj^ii+b)*((xxmat/jj)^ii-1))	
  outer <- sum(inner^2)
  
  y <- outer
  return(y)
}

```

### 6.3) Visulization of Perm $P_{n,\beta}$ Function
```{r}
visualize3D(perm , 2)
```


## 7) Rastrigin Function 
### 7.1) Equation of Rastrigin Function 

$$
f(x) = 10d + \sum_{i=1}^{d}[x_i^2 -10cos(2\pi x_i)]
$$


### 7.2) R implementatin of Rastrigin Function

```{r Rastrigin Function}

rastrigin <- function(x) {
  y <- 10 * length(x) + sum(x^2 - 10 * cos(2 * pi * x))
  return(y)
}

```


### 7.3) Visulization of of Rastrigin Function
```{r Rastrigin Function Visulization}
visualize3D(rastrigin,5)
```

