---
title: "Test 2"
author: "Siddhant Raghuvanshi"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1) Functions to be optimized

The functions defined in this section were created in test 1.

##  1.1) Ackley function

```{r Ackley}
ackley<-function(x,a=20, b=0.2 , c =2*pi){
  d <- length(x)
  
  sum1 <- sum(x^2)
  sum2 <- sum(cos(2*pi*x))
  
  term1 <- -a * exp(-b*sqrt(sum1/d))
  term2 <- -exp(sum2/d)
  
  y <- term1 + term2 + a + exp(1)
  return(y)
}

```

## 1.2) Rastrigin function 

```{r Griewank function}

rastrigin <- function(x) {
  y <- 10 * length(x) + sum(x^2 - 10 * cos(2 * pi * x))
  return(y)
}

```

## 1.3) The Michalewicz function
```{r}
michalewicz <- function(x,m=10){
d <- length(x)
i <- seq_along(x)

y <- sum(sin(x) * (sin(i*x^2/pi))^(2*m))
return(-y)
}
```

# 2) Documentation of steps taken.

## Step 1 - Overview of Optim

I stared with overview of Optim function, its features and how to use it. A simple Google search revealed that it is included in base R with stats package which is loaded by default when an R environment is opened so I ran the command `help("optim")` in Rstudio to get some idea about syntax and parameters.

The built-in R documentation showed that `optim()` is meant to be a general-purpose optimizer which is based on Nelder–Mead, quasi-Newton and conjugate-gradient algorithms. It includes an option for box-constrained optimization and simulated annealing.By default optim calculates the minima of the functions.

## Step 2 -  Understanding Arguments of Optim

1. par - a vector of values which can be passed to the function which is to be optimized.
2. fun - function to be optimized.
3. gr - gradient of the function.
4. method - choice of method to be used for optimization task.
5. lower and upper = lower and upper bound respectively for constrained optimization. 
6. control - a list of parameter to control and track a solver.
7. hessian - returns a hessian if true.

## Step 3 - Understanding Output and Methods

The function returns a list containing the following :-

1.  par - the final value of parameter for which the function achieves optima
2.  value - the optimized value found for the final parameters.
3.  counts - the two values shows number of calls and function and gradient respectively.
4.  convergence - an integer value to show if the algorithm converged or something else happened corresponding to the algorithm used
5.  message - provides some additional information given by the optimizer

## Step 4 - Optimizing functions with optim
'TestMethod()' is a small function that i wrote which takes a function for which we wish to calculate minima and then it evaluates minima using optim along with that it prints a table containing the output. The motivation for this function is to compare various methods at once.
```{r TestMethod}
testMethod <- function(val,fname){
  
  method = c("Nelder-Mead", "BFGS", "CG", "L-BFGS-B", "SANN")
  
  table<-NULL
  cat("Values for point: ",val,"\n\n")
  
  for (i in method){
    opt<- optim(val,fname,method = i)
    parameter<-opt$par
    
    # need to find a better way to pass parameters in cbind!
    row<- cbind(Method=i, Parameter1 = opt$par[1], Parameter2 = opt$par[2],Parameter3 = opt$par[3],Parameter4 =opt$par[4],
                value = opt$value, convergence = opt$convergence )
    table<-rbind(table,row)
  }

  print(table)
  cat("\n\n")
}
```

### 4.1) Optimizing Ackley function
For Ackley function Minimum value is 0.

```{r optim on ackley}



testMethod(c(1,1,1,1),ackley)
testMethod(c(2,2,2,2),ackley)
testMethod(c(3,3,3,3),ackley)


```
### 4.2) Optimizing Rastrigin function
For Rastrigin function Minimum value is 0.
Surprisingly none of the solvers do well for this function and all of the solvers converge.
```{r otim on griewank}
testMethod(c(1,1,1,1),rastrigin)
testMethod(c(2,2,2,2),rastrigin)
testMethod(c(3,3,3,3),rastrigin)
```
### 4.3) Optimizing Michalewicz function
For 5 dimensional input the minima value is -4.6876
```{r michalewicz function}
testMethod(c(1,1,1,1,1),michalewicz)
testMethod(c(2,2,2,2,2),michalewicz)
testMethod(c(3,3,3,3,3),michalewicz)
```
